---
title: "TpitReport"
author: "W. McDonald"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

The goal of this document is to reproducibly import, wrangle and analyse the full pituitary adenoma tissue microarray dataset, including all immunostains up to *Tpit*. *CAM5.2* staining pattern is included, as a relevant categorical variable. Otherwise, IHC staining extent and intensity are reported as Allred scores as previously published. 

This analysis will rely on the tools of the tidyverse, when possible. 

## Data import, wrangling.

```{r echo=TRUE}
library(tidyverse)

patma <- read_csv("data/TpitAddedVillaUpdate.csv")

newClass <- patma %>%
  count(finDx, sort = TRUE)
newClass
sum(newClass$n)

oldClass <- patma %>% 
  count(Dx, sort = TRUE)
oldClass
sum(oldClass$n)

```

Much work went into applying the Villa algorithm on or dataset and setting the finDx as nearly as possible in terms of their nomenclature. A simple comparison of the resulting classes is in order:

```{r}
mean(patma$finDx == patma$villaCode)
```

## Correlation matrix and p values for the correlations

Beyond the R help documentation for cor(), I like the visualizations from [this site that uses corrplot](http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram)

The package Hmisc is used to generate p-values for the correlation matrix.

```{r}
library(corrplot)
patma %>% 
  select(SF1Median, Pit1Median, TPITMedian, PRLMedian, GHMedian, TSHMedian, LHMedian, FSHMedian, ACTHMedian) %>% 
  cor(y = NULL, use = "complete.obs", method = "pearson") %>% 
  round(2)

M <- patma %>% 
  select(SF1Median, Pit1Median, TPITMedian, PRLMedian, GHMedian, TSHMedian, LHMedian, FSHMedian, ACTHMedian, ASUMedian, GATA3Median) %>% 
  cor(y = NULL, use = "complete.obs", method = "pearson")

corrplot(M, type = "upper", order = "hclust")

# mat : is a matrix of data (function by Alboukadel Kassambara
#PhD, see his webpage, cited above)
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(M)
head(p.mat[, 1:5])

# Specialized the insignificant value according to the significant level: this crosses out insig. values
corrplot(M, type="upper", order="hclust", 
         p.mat = p.mat, sig.level = 0.05)

# Leave blank on no significant coefficient
corrplot(M, type="upper", order="hclust", 
         p.mat = p.mat, sig.level = 0.05, insig = "blank")

# Another way to get p-values manually is to use the Hmisc package: 
library(Hmisc)

pvalues <- patma %>% 
  select(SF1Median, Pit1Median, TPITMedian, PRLMedian, GHMedian, TSHMedian, LHMedian, FSHMedian, ACTHMedian, ASUMedian, GATA3Median) %>% 
  cor(y = NULL, use = "complete.obs", method = "pearson") %>%
  as.matrix() %>%
  rcorr()

pvalues

```
