---
title: "managerialDx"
author: "W. McDonald"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Managerial Diagnoses

## Import and clean step

```{r}
library(tidyverse)

patma <- read_csv("data/TpitAddedVillaUpdate.csv")

# Get rid of the Note column
patma <- patma[-30]
glimpse(patma)
```

## Too many classes

Note the composition of finDx:

```{r}
patma %>% count(finDx, sort = TRUE)
```

This is 17 fine-grained diagnostic classes. Too many for a variety of reasons. A managerial group would greatly simplify the task of supervised and unsupervised statistical learning (machine learning) and reduce the likelihood of overfitting. 

## Use mutate() and case_when to generate a new column: manDx

```{r}
x <- patma %>% mutate(manDx = case_when(
  finDx == "GON" ~ "GON",
  finDx %in% c("ACTH(s)", "ACTH(DG)") ~ "ACTH",
  finDx %in% c("PRL", "PRL+") ~ "PRL",
  finDx == "NULL" ~ "NULL",
  finDx %in% c("PlurGH", "GH+(SG)", "GH(SG)", "GH+(DG)", "GH") ~ "GH",
  finDx == "PLUR" ~ "PLUR",
  finDx %in% c("GH/PRL", "PIT1(s)", "TSH(s)", "TSH+") ~ "PIT1",
  TRUE ~ "UNK"
))

x %>% count(manDx, sort = TRUE)
```

