---
title: "reorderingTable"
author: "W. McDonald"
date: "8/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reordering a table for PitAdTMA8.0/TpitAnalysis

The following is based on a a very nicely illustrated post on Stack Overflow at 

<https://stackoverflow.com/questions/42497063/r-sorting-factor-by-level-frequency-and-plotting>

```{r}
library(tidyverse)

set.seed(101)
df <- data.frame(Var1=as.factor(sample(c("AB", "BC", "CD", "DE", "EF"),
                            prob=c(0.1,0.5,0.2,0.05,0.15),
                                   1000, replace=TRUE)))
You want reorder(), I think. In general reorder(x,y,FUN) changes the order of levels of a factor x according to the results of applying function FUN to a second variable y. In this case, you can use FUN=length, and it doesn't really matter what you use for y.

Set up data (I chose to modify the probabilities to make the result clearer):

set.seed(101)
df <- data.frame(Var1=as.factor(sample(c("AB", "BC", "CD", "DE", "EF"),
                            prob=c(0.1,0.5,0.2,0.05,0.15),
                                   1000, replace=TRUE)))

# Wrong order

print(g1 <- ggplot(df, aes(x=factor(1), fill=Var1))+
   geom_bar(width=1, colour="black")+
   coord_polar(theta="y")+
   theme_void())
```

Note the nice bit of sampling for future use.

Now reorder the factors:

```{r}
df$Var1 <- reorder(df$Var1,df$Var1,FUN=length)
levels(df$Var1)
```

Check that the order is correct:

```{r}
sort(table(df$Var1))
```

Now the author of this post does something that loses me:

> "Print the new plot (stick in the new data with %+% and flip the order of the legend with guide_legend(): you could also use function(x) -length(x) as FUN to change the order of the levels in the first place)."

```{r}
print(g1 %+% df + 
    scale_fill_discrete(guide=guide_legend(reverse=TRUE)))
```



